Developing Data Products using R
========================================================

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 160)
```

```{r}
rm(list=ls())
```


AppliedPredictiveModeling package is used to reproduce the analyses in the text. 

```{r}
library(caret)
library(AppliedPredictiveModeling)
```

```{r}
data(concrete)
```

Concrete - Data sets from different sources that can be used for modeling the compressive strength of concrete formulations as a functions of their ingredients and age.
```{r}
#?concrete
```

```{r}
summary(concrete)
```

```{r}
names(concrete)
```

```{r}
dim(concrete)
```

Check total missing values 
```{r}
sum(is.na(concrete))
```

Partition the dataset
```{r}
set.seed(975)
inTrain = createDataPartition(mixtures$CompressiveStrength, p = 3/4)[[1]]
```

into test and traininig datasets
```{r}
training = mixtures[ inTrain,]
testing = mixtures[-inTrain,]
```

```{r}
head(training)
```

```{r}
str(training)
```

```{r}
#str(concrete)
```

```{r}
qplot(Superplasticizer, data=concrete)
```

```{r}
qplot(log(Superplasticizer + 2), data=concrete)
```

The skewness cannot be removed, becuase the values are exactly the same. 


```{r}
set.seed(3433)
data(AlzheimerDisease)
```

```{r}
names(predictors)
```

```{r}
head(diagnosis)
```

Factor with two levels - Impaired and Control
```{r}
attributes(diagnosis)
```

```{r}
adData = data.frame(diagnosis,predictors);names(adData)
```

```{r}
dim(adData)
```

```{r}
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]
```

Select the names which start with IL only
```{r}
tr  <- training[colnames(training)[grepl('^IL',colnames(training))]]
summary(tr)
```

Using pca recuve the number of features to account for 80% variance. 
```{r}
preProc  <- preProcess(tr,method="pca",thresh = 0.8)
preProc
```

```{r}
rm(list=ls())
set.seed(3433)
library(AppliedPredictiveModeling)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]
```

```{r}
train_il  <- training[colnames(training)[grepl('^IL',colnames(training))]]
str(train_il)
```

Prediction function # 1 
```{r}
library(caret)
head(diagnosis[inTrain])
length(diagnosis[inTrain])
```

glm - Generalized Linear Models 
Fit the dataset using `glm`. 
```{r}
modelFit  <- train(diagnosis[inTrain] ~ ., method="glm",data=train_il)

ab  <- cbind(train_il,diagnosis[inTrain])
ab$diagnosis[inTrain]
modelFit  <- train(ab$diagnosis[inTrain] ~ ., method="glm",data=ab)
```


Using Plotly 
------------

Initialize a Plotly object :
```{r}
library(plotly)
py <- plotly()
```

Now you can make interactive, online graphs with `ggplot2` syntax using `py$ggplotly()`


```{r Example_1}
(ggiris <- qplot(Petal.Width, Sepal.Length, data = iris, color = Species))

r <- py$ggplotly(ggiris)
r$response$url
```



```{r Exmaple_2}
library(plotly)

dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
d <- qplot(carat, price, data=dsamp, colour=clarity);d

py$ggplotly()
```



Manipulate
----------


```{r Manipulate}
library(manipulate)
library(HistData)
data(Galton)
myHist <- function(mu){
  hist(Galton$child,col="blue",breaks=100)
  lines(c(mu, mu), c(0, 150),col="red",lwd=5)
  mse <- mean((Galton$child - mu)^2)
  text(63, 150, paste("mu = ", mu))
  text(63, 140, paste("MSE = ", round(mse, 2)))
}
manipulate(myHist(mu), mu = slider(62, 74, step = 0.5))

```


```{r Manipulate 2}
library(manipulate)
myPlot <- function(s) {
    plot(cars$dist - mean(cars$dist), cars$speed - mean(cars$speed))
    abline(0, s)
}


manipulate(myPlot(s), s = slider(0, 2, step = 0.1))
```






rCharts
--------
```{r nplot}
require(rCharts)
haireye  <- as.data.frame(HairEyeColor)

n1  <- nPlot(Freq ~ Hair, group = 'Eye', type = 'multiBarChart',
             data = subset(haireye, Sex == 'Male'))

n1$save('figure/n1.html', cdn = TRUE)
cat('<iframe src="figure/n1.html" width=100%, height=600></iframe>')
```


```{r}
nPlot(Freq ~ Hair, group = 'Eye', type = 'multiBarChart',
             data = subset(haireye, Sex == 'Male'))
```

bring up the plot 
```{r}
n1
```

Displays the html required for the plot 
```{r}
n1$html()
```


Another example 
```{r Facetted Scatterplot}
names(iris) = gsub("\\.","",names(iris))
r1  <- rPlot(SepalLength ~ SepalWidth | Species, data = iris, 
             color = "Species", type = "point")
r1$save("figure/r1.html", cdn = TRUE)
cat('<iframe src="figure/r1.html" width=100%, height=600></iframe>')
```


```{r Facetter Scatterplot}
hair_eye  <- as.data.frame(HairEyeColor)
r2  <-rPlot(Freq ~ Hair| Eye, color = "eye", data = hair_eye, type="bar")
r2
r2$save("figure/r2.html", cdn = TRUE)
cat('<iframe src = "figure/r2.html" width=100%, height=600></iframe>')
```


How to get the js, html or publish an rChart 

```{r}
r1  <- rPlot(mpg ~ wt| am + vs, data = mtcars, type="point", color = "gear"); r1
r1$print("chart1") # Print out the js 
r1$save('myPlot.html') # Save the html file.
# ------
# not working 
r1$publish('myPlot.html', host='gist') # save to gist, rjson required 
r1$publish('myPlot', host = 'rpubs') # Save ti rpubs
```


```{r}
r1$publish
```

```{r morris}
data(economics, package="ggplot2")
?economics
str(economics)
econ  <- transform(economics, data = as.character(date))
str(econ)

m1  <- mPlot(x = "date", y = c("psavert", "uempmed"), type="Line", data = econ); m1
m1$set(pointSize=0, lineWidth=1);m1

```



```{r xCharts}
require(reshape2)
uspexp  <- melt(USPersonalExpenditure)
View(USPersonalExpenditure)
View(uspexp)
names(uspexp)[1:2]  <- c("category", "year")
x1  <- xPlot(value ~ year, group = "category", data = uspexp, type="line-dotted");x1
```




```{r Leaflet}
map3  <- Leaflet$new()
map3$setView(c(51.505, -0.09), zoom = 13)
map3$marker(c(51.5, -0.09), bindPopup = "<p> Hi, I am a popup </p>")
map3$marker(c(51.495, -0.083), bindPopup = "<p> Hi, I am another popup </p>")
map3
```


```{r Rikshaw}
usp  <- reshape2::melt(USPersonalExpenditure)
# Get the decades into a date Rickshaw likes 
usp$Var2  <- as.numeric(as.POSIXct(paste0(usp$Var2, "-01-01")))
p4  <- Rickshaw$new()
p4$layer(value ~ Var2, group = "Var1", data = usp, type = "area", width = 560); p4

# Add a helpful slider this easily; other features TRUE as default
p4$set(slider=TRUE); p4
```




```{r highchart}

h1  <- hPlot(x = "Wr.Hnd", y = "NW.Hnd", data = MASS::survey, 
             type=c("line","bubble","scatter"), group = "NW.Hnd",
             size="Age");h1

library("MASS")
?survey # Student Survey Data
#Wr.Hnd: span (distance from tip of thumb to tip of little finger of spread hand) of writing hand, in centimetres.
#NW.Hnd:span of non-writing hand.


dataSet  <- MASS::survey; View(dataSet[,c("Wr.Hnd", "NW.Hnd", "Age")])

```


```{r}
require(rCharts)
data(airquality)
View(airquality)
dTable(airquality, sPaginationType = "full_numbers")
```




Shiny 

```{r}
library(shiny)
```


